### 2011-04-26 papers/fse2011/reviewResponse.txt
edited version of Salman's response

Thank you for reviewing our paper on 'lastChange'. We respond to the
questions raised by the reviewers and then request that the first
reviewer reconsider part of their analysis.

The first reviewer asks:

Q: "The paper presents the work as a general debugging approach. ... how
much of it is made possible through the unique characteristics of
Firefox and Javascript programming? How closely similar could it be
possible to support this idea for a compiled language?"

A: Section 3 of our paper presents a general algorithm and in that
section we give the four prerequisite operations, 1) objectId(), 2)
setPropertyChangeHook(), 3) scopeId(), and 4)
setVariableChangeHook(). A runtime which can support these operations
can run the algorithm; other runtimes cannot. No special feature of
Firefox or of JavaScript is used in this section. Whether a language
is compiled or not does not relate to our technique. We have
implemented this technique for Java as shown in an unreviewed
technical report: http://infoscience.epfl.ch/record/164417. This
technique should apply to C, C++, C#, Pascal, PHP, Ruby, and so on.

The second reviewer asked many important questions about our small user study:

Q: "Who were the developers?"

A: Unfortunately it would not be practical to give the names of our study
particiapants. We assume the reviewer is asking for more information
about the skill of the developers, which we addresss in the remaining
answers.

Q: "How much experience did they have?"

A: These were all skilled developers with several years of experience in
debugging complex programs.

Q: "Are they researchers?"

A: Yes of course. Practically speaking, we cannot hire 'average'
developers to engage in a lengthy study of such a complex subject as
debugging code. A small study with "folks down the hall" is what is
practical.

Q: "Do they know the concept of a data dependency?"

A: Yes. However, our prototype does not require users to know this concept.

Q: "Do they work on programming tools?"

A: No.

Q: "How complex were the test applications?"

A: These were simple applications and they are available on the Web site
we listed in footnote 4,
http://ltiwww.epfl.ch/~mirghase/lastchange-userstudy.

Q: "We're they already familiar with Firebug and its debugging tools?"

A: Of course, most JavaScript developers use Firebug.

Q: "How was the task explained to them?"

A: The instructions were provided on the Web pages listed in footnote 4,
http://ltiwww.epfl.ch/~mirghase/lastchange-userstudy/group1.html.

Q: "We're they given time limits?"

A: As described in the Web pages listed in footnote 4, the introduction
and practice sessions where not timed, then two time trials were
performed. That is how we arrived that the numbers in Figure 12.

Q: "How were they recruited?"

A: We asked for volunteers from among people in our building who we
knew wrote JavaScript.

Q: "What counted as a "step" in Figure 12)."

A: Oops we should have explained that a step in Figure 12 means a button
push, either single stepping the debugger or running a querypoint.

Q: "The sample size of 4 was extremely small, preventing the authors from
really avoiding learning effects (they did counterbalance exposure to
each program)."

A: We did counterbalance amoung users but as you say the small sample
makes this ineffective. We would like to point out that learning
effects stongly bias against our technique. As the reviewers surely
experience, debugging takes a long time to master. New debugging
techniques take a long time to learn. That our test subjects were able
to succeed in less than an hour is itself the most important result of
our small study.

Finally we ask reviewer #1 to reconsider section 4 of our paper and
the related work section discussing reverse watchpoints.

We are very surprised at the claim of reviewer #1 that we don't build
on top of existing debug technology. Section 4 of our paper details
how we built lastChange on top of the Firebug debugger. Perhaps we
should have been clearer: by "breakpoint technology" we mean all the
different ways debuggers pause execution when a special event
occurs. It includes line step, function enter/exit, variable/property
access/change (watchpoint), error/exception, conditional breakpionts,
and changes on graphical changes. (see for example our ref. 1). Our
ability to build upon existing debugger technology is essential to the
practicality of our approach.

We do not believe the claim of the first reviewer that a developer can
"manually perform a lastChange query in relatively few steps." In fact
our user study, as small and limited as it may be, provides direct
evidence against this claim. Moreover our approach can be applied
recursively, to obtain the lastChange to a value found on a
lastChange.

Reviewer #1 points to a combination of watchpoints and the "drop in
frame" feature in Eclipse. Watchpoints trigger on data value changes;
"Drop in frame" allows for shorter re-execution cycles like the local
replay we explained in section 4.5. The fundamental problem with these
techniques is that they only work in the 'forward' direction. If we
land on a watchpoint after a 'drop in frame' operation, then is this
the last change or just one of a hundred changes? This fundamental
problem is why papers cited as our references 4, 5, 6, 8, 11, 12, 13,
and 15 exist: developers want to go backwards. That is why we called
our paper "lastChange". As cited in our related work section, our
approach is superior to even reverse watchpoints, a technology created
to go beyond normal watchpoints. We could of course include a
discussion of normal watchpoints.

Again we thank the reviewers. We believe small changes to the text
based on this feedback would clarify and improve the paper.