### 2011-04-27 papers/fse2011/reviewResponse.txt

Thank you for reviewing our paper on 'lastChange'. We respond to the
questions raised by the reviewers and then request that the first
reviewer reconsider part of their analysis.

The first reviewer asks:

Q: "The paper presents the work as a general debugging approach. ...  
How closely similar could it be possible to support this idea for a compiled language?"

A: Section 3 of our paper presents a general algorithm and we have
implemented this technique for Java as shown in an unreviewed
technical report: http://infoscience.epfl.ch/record/164417.

The second reviewer asked many important questions about our small user study, we summarized them in 3 main qeustions:

Q: Who were the developers?

A: The developers were all skilled developers with several years of experience in
debugging complex programs and familiar with Firebug. They knew the concept of 
data dependency, however our prototype does not require users to know this concept. 
We recruited them by asking for volunteers from among people in our building who we 
knew wrote JavaScript.

Q: How the user study is conducted?

A: The instructions of user study and programs code were provided on the Web site we listed in footnote 4,
http://ltiwww.epfl.ch/~mirghase/lastchange-userstudy .

Q: What counted as a "step" in Figure 12?

A: We should have explained that a step in Figure 12 means a button
push, either single stepping the debugger or running a lastChange query.

We are very surprised at the claim of reviewer #1 that we don't build
on top of existing debug technology. Section 4 of our paper details
how we built lastChange on top of the Firebug debugger. Perhaps we
should have been clearer: by "breakpoint technology" we mean all the
different ways debuggers pause execution when a special event
occurs. It includes line step, function enter/exit, variable/property
access/change (watchpoint), error/exception, conditional breakpionts, etc.
Our ability to build upon existing debugger technology is essential to the 
practicality of our approach.

Reviewer #1 points to a combination of watchpoints and the "drop in
frame" feature in Eclipse. Watchpoints trigger on data value changes;
"Drop in frame" allows for shorter re-execution cycles like the local
replay we explained in section 4.5. The fundamental problem with these
techniques is that they only work in the 'forward' direction. If we
land on a watchpoint after a 'drop in frame' operation, then is this
the last change or just one of a hundred changes? This fundamental
problem is why papers cited as our references 4, 5, 6, 8, 11, 12, 13,
and 15 exist: developers want to go backwards. That is why we called
our paper "lastChange". As cited in section 8, our
approach is superior to even reverse watchpoints, a technology created
to go beyond normal watchpoints. We could of course include a
discussion of normal watchpoints.

Again we thank the reviewers. We believe small changes to the text
based on this feedback would clarify and improve the paper.

--------------------------------------------------------------------------------
Removed
--------------------------------------------------------------------------------
We do not believe the claim of the first reviewer that "a developer can 
manually perform a lastChange query in relatively few steps". In fact
our user study, as small and limited as it may be, provides direct
evidence against this claim. Moreover our approach can be applied
recursively, to obtain the lastChange to a value found on a
lastChange.